---
title: "HW3 Peer Assessment"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

For 30 Galapagos Islands, the data consist of the number of plant species found on each island and the number that are endemic to that island. We also have five geographic variables for each island. The data was presented by Johnson and Raven (1973) and also appear in Weisberg (2005).

While the dataset contains seven variables, we are interested only in  the following six:

Species: the number of plant species found on the island
Area: the area of the island (km^2)
Elevation: the highest elevation of the island (m)
Nearest: the distance from the nearest island (km)
Scruz: the distance from Santa Cruz island (km)
Adjacent: the area of the adjacent island (km^2)

Please install the package faraway, load the dataset gala, and drop the unnecessary second column that is Endemics:

## Read the data

```{r}
require(faraway)

```


# Question 1:  Fitting the Model - 10 pts

*Fit a poisson regression model using the number of species as the response variable and all other variables as predictors. Call it model1.*

**(a) 3 pts - Display the summary of model1. What are the model parameters and estimates?**

```{r}
# Your code here...

```


**(b) 3 pts - Write down the equation for the estimated number of species given the predicting variables.**


**(c) 4 pts - Provide meaningful interpretations for the coefficients of Elevation and Scruz.**


# Question 2: Inference - 9 pts

**(a) 3 pts - Using model1, find a 90% confidence interval for the coefficient for Nearest.**

```{r}
# Your code here...

```


**(b) 3 pts - Is model1 significant overall? How do you come to your conclusion?**

```{r}
# Your code here...

```


**(c) 3 pts - Which variables are significantly nonzero at the 0.01 significance level? Which are significantly positive? Why?**


# Question 3: Goodness of fit - 12 pts

**(a) 4 pts - Perform goodness of fit hypothesis tests using both deviance and Pearson residuals. What do you conclude? Explain the differences, if any, between these findings and what you found in Question 2b.**

```{r}
# Your code here...

```


**(b) 5 pts - Perform visual analytics for checking goodness of fit for this model and write your observations. Be sure to address the model assumptions. Only deviance residuals are required for this question.**

```{r}
# Your code here...

```


**(c) 3 pts - Calculate the dispersion parameter for this model. Is this an overdispersed model?**

```{r}
# Your code here...

```


# Question 4: Fitting a Count per Area Model - 21 pts

*Let's create a rate based poisson regression model for the same dataset. Now the response will be density of species (number of species/km^2). So the exposure in this case will be Area. Call this model2. Fit the model and display the summary of the model.*

**(a) 2 pts - Write down the equation for the estimated number of species per square kilometer given the predicting variables.**

```{r}
# Your code here...

```


**(b) 2 pts - Provide a meaningful interpretation for the coefficient of Adjacent.**


**(c) 2 pts - Is Nearest significant given the other variables?**


**(d) 12 pts - Has your goodness of fit been affected? Repeat the tests, plots, and dispersion parameter calculation you performed in Question 3 with model2.**

```{r}
# Your code here...

```


```{r}
# Your code here...

```


```{r}
# Your code here...

```


**(e) 3 pts - Overall, would you say model2 is a good-fitting model? If so, why? If not, what would you suggest to improve the fit and why? Note, we are not asking you to spend hours finding the best possible model but to offer plausible suggestions along with your reasoning.**

```{r}
library(mgcv)
```


# Question 5: Prediction - 6 pts

*Suppose you've found a new island 110 km from Santa Cruz with an area of 25 km^2 and an elevation of 100 m. Its nearest island is 21.1 km away with an area of 0.57 km^2.*

**(a) 2 pts - Predict the number of plant species on this new island using model1.**

```{r}
# Your code here...

```


**(b) 2 pts - Predict the number of plant species on this new island using model2.**

```{r}
# Your code here...

```


**(c) 2 pts - Comment on how your predictions compare.**
